### 2018-02-23 thesis/contents/tn_sagit.tex

\begin{table}[ht]
\centering
\csvautotabular{tn_sagit_table.csv}
\caption{List of the best GP models for different diffusivity measures and window segments. Only models with at least 70\% accuracy were accepted. Precision and f1 scores are also provided for reference}
\label{table:GP}
\end{table}

\section{Discussion}
The study is the first instance of machine learning discrimination of TN vs. Controls with specific white matter analysis, using group-wise merged along-the-track tractography. In essence, we have demonstrated a fully-automated, end-to-end tractography and machine learning platform that is capable of discriminate TN from white-matter diffusivity with greater than 80\% accuracy solely from white-matter measures. We reproduced previous findings of TN diffusivity changes along the CN V with much higher regional specificity. We also examined TPT and S1 projections of the trigeminal sensory pathway.

The diffusivity measures of CN V reproduced our previous findings \cite{Chen2016a}, which was that cistern, REZ differentiated TN from controls. The consistent regional overlaps in the cistern/ganglion region of the right (affected) CN V showed the ability for the GP classifiers to auto-converge to focal white-matter abnormality. The left versus right GP classifier placed left-right differences at the cistern/REZ segment, in agreement with the pathophysiology of TN. The regional findings on left versus controls suggest that there is bilateral differences in CN V myelin in unilateral TN, despite the lack of symptoms on the unaffected side. The left regions also situate into the brainstem segments. It is possible that the left CN V is affected by global descending pain modulation, and warrant further investigation.

The TPT delineations showed that decussation pathway tractography is possibly asymmetric. It demonstrated very strong differences from the classifier results, with overlapping differences in the ascending segment. There also seemed to be diffusivity differences in the decussation segments between affected and unaffected sides. The decussaing fibers of the TPT emerges from the TGN, therefore the difference seemed reasonable. It is not clear however, whether the tractography asymmetry of the TPT is neuroanatomically factual, or an artifact from either algorithmic bias or partial volume bias of the underlying DWI. Resolving crossing fiber is still a difficult task in tractography, and while there are validations against MR phantoms, the anatomical validity of these methods can only be extrapolated, and the interpretation must involve anatomical priors from experts.

>>>>

\begin{table}[ht]
\centering
\csvautotabular{tn_sagit_table.csv}
\caption{List of the best GP models for different diffusivity metrics and centroid windows. Only models with at least 70\% accuracy were accepted. Table is sorted by anatomical region, classifier type and accuracy. Start and end denote the beginning and end of the window subdivision positions along the centroid. Length is the number of subdivisions covered by the window. Precision and f1 scores are provided for reference}
\label{table:GP}
\end{table}

\section{Discussion}
We have demonstrated the use of a fully-automated, end-to-end tractography and machine learning platform that is capable of discriminate TN from white-matter diffusivity with greater than 80\% accuracy strictly from white-matter measures. Our work is in agreement with previous findings of TN diffusivity changes along the CN V with much higher regional specificity. We also examined TPT and S1 projections of the trigeminal sensory pathway. The study is the first instance of machine learning discrimination of TN vs. Controls with specific white matter analysis, using group-wise merged along-the-track tractography.

The diffusivity measures of CN V agreed with our previous findings \cite{Chen2016a}, which was that cistern, REZ differentiated TN from controls. The consistent regional overlaps in the cistern/ganglion region of the right (affected) CN V showed the ability for the GP classifiers to auto-converge to focal white-matter abnormality. The GP classifier placed affected vs unaffected differences at the cistern/REZ segment, in agreement with the pathophysiology of TN. The findings on unaffected versus controls suggest that there is bilateral differences in CN V myelin in unilateral TN, despite the lack of symptoms on the unaffected side. The left regions also situate into the brainstem segments. It is possible that the left CN V is affected by global descending pain modulation, and warrant further investigation.

The TPT delineations showed that decussation pathway tractography is possibly asymmetric. It demonstrated very strong differences from the classifier results, with overlapping differences in the ascending segment. There also seemed to be diffusivity differences in the decussation segments between affected and unaffected sides. The decussating fibers of the TPT emerges from the TGN, therefore the difference seemed reasonable. It is not clear however, whether the tractography asymmetry of the TPT is neuroanatomically factual, or an artifact from either algorithmic bias or partial volume bias of the underlying DWI. Resolving crossing fiber is still a difficult task in tractography, and while there are validations against MR phantoms, the anatomical validity of these methods can only be extrapolated, and the interpretation must involve anatomical priors from experts.
